---
title: "Lab 6 loops and functions"
author: "Amelia Ritger"
date: "11/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Why functions?

- They make it easy to do a task over and over without copy-and-pastng the code.
- You can take a complicated piece of code and give it a useful descriptive name.

## Making a basic function

Make a function to take a number and add 1
```{r}
add_one <- function(x) {
  y <- x+1
}

z <- add_one(x=4)
zz <- add_one(3:7)
```

```{r}
add_some <- function(x,a=1) {
  y <- x+a
  print("yay!")
  return(y)
}

add_some(x=4)
```

## Make a useful function!

A good practice is to use verb names for your functions.

Let's make a function to calculate the mean of a vector of values. First let's create a test vector of values we can use for testing. We'll use `rnorm()` to generate a pseudo-random set of numbers as a working sample, with a known mean and sd.
```{r}
set.seed(12345)
#'12345' can be any number, but as long as we use the same number, we should get the same sample
#because with set.seed() you are telling R where to start along a sequence of set "random" numbers (aka pseudo-random)
#set.seed() for perfect reproducibility of your results
test_vec <- rnorm(n=100, mean=1, sd=1)
head(test_vec)
mean(test_vec)
sd(test_vec)

calc_mean <- function(x) {
  n <- length(x)
  my_mean <- sum(x)/n
  return(my_mean)
}

calc_mean
calc_mean(test_vec) #woot. same as mean(test_vec) value
```

## Anatomy of a function:

- `function()` is a function that defines functions
- Inside the arguments for `function(...)` are the information we need to give the function for it do to its job.
- After the arguments, inside the {}, we put the code that defines what the function does.
- At the end, the `return()` gives back to the user the desired output of the function.

## Environments

Most of the time we're working in the Global Environment.

When you execute a function, R starts a new independent environment, like a pocket universe, separate from the global environment. The arguments (in this case, x) allow a window through which we can pass information to that new enviornemnt; the return() is a window to  pass information back.
```{r}
x <- 100
y <- 200
z <- add_one(x=8)
```

Note that changes in variables inside the function are NOT communicated back to the gobal environment. That's a good thing!

A well-writen function should communicate with the global environment only through arguments and return values.

One you're done running a function, that new enviornment returns its value then disappears forever.

## Improving our basic function

A well-written function should anticipate whhen the user might have imperfect data. Error checks or additional arguments can be helpful. When might our calc_mean() function break down?
```{r}
calc_mean(1:10)
calc_mean(x=1:10)
calc_mean(x=c(1,2,3,4,NA)) #whenever an NA is present the function breaks
```

## Let's create a test vector with some NAs

Also, let's learn abuot indexing to access specific elements in a vector. Use "index" numbers to tell which elements of the vector to change. 
```{r}
test_vec_w_NA <- test_vec #just to make a copy
test_vec_w_NA[5] <- NA #make element #5 an NA
test_vec_w_NA[10:15] <- NA #likewise...
test_vec_w_NA[c(20,22,24)] <- NA 
```

What do we need to do to make our function work with NAs?

- Design in an na.rm argument to allow the user to decide whether to exclude NAs.
- Design in a way to exclude NAs if the user asks.

### Another quick thing with vectors:

How to exclude NAs?
```{r}
z <- c(1:5, NA)
is.na(z)
zz <- z[!is.na(z)]

calc_mean2 <- function(x_vec, na.rm=TRUE) {
  if(na.rm==TRUE) { #double equal sign is a logical test, single equal sign is an assignnment
    x_vec <- x_vec[!is.na(x_vec)]
  }
  n <- length(x_vec) # make user assign a vector, indicate that function should be expecting a vector
  mymean <- sum(x_vec)/n
  return(mymean)
}

calc_mean2(test_vec_w_NA, na.rm=TRUE)

## and here's a proof of concept 4fun
calc_mean2 <- function(x_vec, no_more_na="apples") {
  if(no_more_na=="pears") { #double equal sign is a logical test, single equal sign is an assignnment
    x_vec <- x_vec[!is.na(x_vec)]
  }
  n <- length(x_vec) # make user assign a vector, indicate that function should be expecting a vector
  mymean <- sum(x_vec)/n
  return(mymean)
}

calc_mean2(test_vec_w_NA, no_more_na="blueberries") #R is looking for "pears" to remove NAs, so because it doesn't see that it moves on from the no_more_na argument
```

```{r}
library(tidyverse)
data(iris)
iris_mean <- iris %>%
  janitor::clean_names() %>%
  group_by(species) %>%
  summarize(mean_sepal_w = calc_mean2(sepal_width))
```

# Why loops?

Sometimes you want to do a thing many times, i..e iterate. Loops are one way to do this, especially if you have a cevtor of values and you want to calculate something based on each value. 
```{r}
times <- 10

for(i in 1:10) {
  print(i+1)
}

for(j in 1:times) {
  x <- x + j
  print(x)
}

for(k in 1:times) {
  print("hello world!")
}
```

A standard die roll has equal chances to land on each of its six sides. Use sample() to simulate this.
```{r}
die_rolls <- data.frame(rolls=sample(x=1:6, size=1000, replace=TRUE))

ggplot(data.frame(die_rolls), aes(x=rolls)) +
  geom_histogram(bins=6)
```

OK, now let's use a loop to make a bunch of simulations, each of 20 rolls, and calculate the mean of each.
```{r}
sims <- 100000
mean_results <- vector('numeric', length=sims)
for(i in 1:sims){
  rolls_vec <- sample(1:6, size=20, replace=TRUE) 
  rolls_mean <- calc_mean2(rolls_vec)
  mean_results[i] <- rolls_mean #"at position i results vector, store mean of rolls_vec"
}

results_df <- data.frame(sim=1:sims,
                         mean=mean_results)

ggplot(results_df, aes(x=mean)) +
  geom_histogram(alpha=0.8, fill='darkred') +
  theme_minimal()

```

### If die rolls are uniformly distributed, why do we see a normal distribution here?

## Central Limit Theorem (yay)!

